---
title: "Fingertips Data Exploration"
format: gfm
---

```{r}
#| eval: false
remotes::install_github("rOpenSci/fingertipsR",
                        build_vignettes = TRUE,
                        dependencies = "suggests")
```


```{r}
library(fingertipsR)
library(kableExtra)
library(tidyverse)
```


Downloading profile list
```{r}
profile_lst <- profiles()
head(profile_lst)
```

Indicator list for Asthma domain (code `8000009`)
```{r}
indicator_lst_asthma <- indicators(DomainID = 8000009)

indicator_lst_asthma |>
  kable()
```

Indicator list for Respiratory disease (code `29`)
```{r}
indicator_lst_resp <- indicators(ProfileID = 29)

indicator_lst_resp |> 
  head(10) |> 
  kable()
```

Selecting the codes for Respiratory profile 

```{r}
sel_indicator <- indicator_lst_resp |>
  pull(IndicatorID) 
```

Indicators are reported for different area types. The next code extracts the area types for each indicator.

```{r}
area_avail_ind <- do.call(bind_rows,lapply(sel_indicator,\(x) indicator_areatypes(IndicatorID = x)))
```

```{r}
area_ty_lst <- area_types()
```


```{r}
area_avail_ind_names <- area_avail_ind |>
  left_join(
    indicator_lst_resp |>
      select(IndicatorID, IndicatorName) |>
      unique(),
    by = join_by(IndicatorID)
  ) |> 
  left_join(
    area_ty_lst |> 
      select(AreaTypeID,AreaTypeName) |> 
      unique(),
    by=join_by(AreaTypeID)
  )

slice_sample(area_avail_ind_names,n = 15) |>
  select(-IndicatorID,-AreaTypeID) |>
  kable()
```


Data for one indicator of the Respiratory profile 
```{r}
temp_data <- fingertips_data(
  IndicatorID = area_avail_ind$IndicatorID[1],
  AreaTypeID = area_avail_ind$AreaTypeID[1]
  )

```


```{r}
# Code from vignette 
cols <- c("IndicatorID", "AreaCode", "ParentName", "Sex", "Timeperiod", "Value")

area_type_name <- table(temp_data$AreaType) # tally each group in the AreaType field

area_type_name <- area_type_name[area_type_name == max(area_type_name)] # pick the group with the highest frequency
area_type_name <- names(area_type_name) # retrieve the name

set.seed(1234)
samp_areas <- sample(temp_data$AreaCode |> unique(),5)


data <- temp_data[temp_data$AreaType == area_type_name & temp_data$AreaCode %in% samp_areas,
                  cols]
```

We are going to show the historic data for five areas of type `r area_type_name`

```{r}
ggplot(data,aes(x = Timeperiod,y = Value, col = AreaCode ,group = AreaCode))+
  geom_line()+
  labs(title = area_avail_ind_names$IndicatorName[1],subtitle = paste0("Showning a sample of ", area_type_name))+
  theme(axis.text.x = element_text(angle = 90))
```

